Sketch Recognition using Enhanced Particle Swarm Optimization

### 2010-10-27 trunk/Swarm/swarmModified.tex
after elvasire editing and tradional pso

\documentclass[10pt]{article}
\usepackage{spconf}
\usepackage{amsmath}
\usepackage[dvips,pdftex]{graphicx}
\usepackage{named}
\usepackage[dvips,pdftex]{hyperref}
\usepackage{subfigure}
\usepackage{array}

\title{Enhanced Particle Swarm Optimization In Sketch Recognition}
\twoauthors
{Maha El Meseery\sthanks{Corresponding Author}, Mahmoud Fakhr El Din}
	{Signals Processing Group \\ Computers and Systems Department \\
  Electronic Research Institute\\Cairo, Egypt\\
	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
  {Tarek abouldahab} {Cairo Metro Company \\
  Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

\begin{document}
\maketitle
 \begin{abstract}
%This study

Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm. The enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols. Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.

\end{abstract}
\pagenumbering{arabic}
 \section{Introduction}

@@

\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}. The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. The following section provides a detailed description of each step. The main block of the system remains the same but their is some modification in the segmentation algorithm to improve its efficiency. The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms. Other modifications are explained in the details in the next sections.  
\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

@@

As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error cased by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}

@@

To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set. The second experiment was tested on both algorithm but only using half the feature set. Both experiments were compared with the original sketch recognition system presented in \cite{mypaper}.

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The figures compares between the result that can be obtained by system in \cite{mypaper} and modifying the system using Enhanced PSO. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.   
    \begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment92.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of the full feature set on both the original system presented in \cite{mypaper} and Enhanced PSO. The results shows that Enhanced PSO increases the performance over the original system. }
   \label{exp2}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment68.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of half the feature set on both the original system presented in \cite{mypaper} and Enhanced PSO. The results shows that Enhanced PSO increase the recognition rate. }
   \label{exp1}
\end{figure}

@@

This paper used an Enhanced-PSO  to modify a new sketch recognition using DPSO represented in \cite{mypaper}. Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features. The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset. Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy. The results proved that the Enhancement made on the PSO improves the efficiency of the Basic PSO algorithm without being dependable on special objective function or application. The system achieved an average overall improvement of more than 1\% over Basic PSO.

@@

\bibliographystyle{IEEEbib}
\bibliography{../../../neededfiles/Bibliographies/Mybibliography}

>>>>

\documentclass[12pt,1p,times]{elsarticle}%[10pt]{article}
%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%\usepackage{spconf}
\usepackage{amsmath}
\usepackage[dvips,pdftex]{graphicx}
%\usepackage{named}
\usepackage{hyperref}
%\usepackage[dvips,pdftex]{hyperref}
\usepackage{subfigure}
\usepackage{array}
\usepackage{amssymb}
%\usepackage{epsfig}
%\usepackage{epstopdf}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}
\begin{document}

\begin{frontmatter}
%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\title{Sketch Recognition using Enhanced Particle Swarm Optimization}

\author[Eri]{Maha El Meseery \corref{cor1} }%,
\ead{melmeseery@eri.sci.eg}
\author[Eri]{Mahmoud Fakhr El Din}
\ead{ mafakhr@mcit.gov.eg}
\author[metro]{Tarek abouldahab}
\ead{heshoaboda@hotmail.com}

\address[Eri]{Signals Processing Group, Computers and Systems Department,
  Electronic Research Institute,Cairo, Egypt.}
  \address[metro]{Cairo Metro Company, Ministry of transport ,Cairo, Egypt.}
\cortext[cor1]{ Corresponding author}

% \twoauthors
% {\sthanks{Corresponding Author},}
% 	{Signals Processing Group \\ Computers and Systems Department \\
%   Electronic Research Institute\\Cairo, Egypt\\
% 	melmeseery@eri.sci.eg, mafakhr@mcit.gov.eg}
%   {} {Cairo Metro Company \\
%   Ministry of transport \\Cairo, Egypt\\ heshoaboda@hotmail.com}

%\maketitle
 \begin{abstract}
%This study

Sketches and drawings are widely used as a simple method of expressing thought and designs. The goal of this paper is to improve our sketch recognition system by using an enhanced particle swarm optimization algorithm. The Enhanced Particle Swarm Optimization (EPSO) is based on the perturbation of the particle velocity is used to optimally segment the strokes the user draws into meaningful geometric primitives.  These geometric primitives are grouped to formulate symbols which are further identified using a Support Vector Machines (SVM) classifier. Experiments were conducted on the benchmark dataset of simple presentation (Hs-DS) symbols. Results show that that the enhanced particle swarm optimization improves the final recognition accuracy.

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
Sketch Recognition, Particle Swarm Optimization
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}
\end{frontmatter}

%\pagenumbering{arabic}
 \section{Introduction}

@@

\section{System Discribtion  }\label{Sysdisc}
%\label{sec:application}
The modified swarm algorithm is used to improve the performance of the sketch recognition system in \cite{mypaper}. The sketch recognitions system is divided into three main steps 1) preprocessing, 2) segmentation and 3) Recognition. Figure \ref{fig:flowChart} shows the flow chart of the system; the first process is collecting stroke points then computing the curvature and possible dominant points.  The following section provides a detailed description of each step. The main block of the system remains the same but there are some modifications in the segmentation algorithm to improve its efficiency. The main modification is using Enhanced Particle swarm algorithm instead of Basic Particle swarm algorithms. Other modifications are explained in the details in the next sections.  
 \begin{figure}
	\centering
		\includegraphics[width=3.5in,height=6in]{FlowChart.jpg}
	\caption{System flow chart} The chart shows the flow of process and data processed thought the system. 
	\label{fig:flowChart}
\end{figure}

\subsection{Preprocessing}
\label{Prepross}
 Speed and time difference information were widely used in sketch understanding systems \cite{earlyprocess}. It was noted that time difference provides more distinct maxima than the speed information. Agar \cite{polygonfeedback31} mentions that the pointing device (i.e. the mouse) sampling rate is the reason for this phenomenon. 
 In the presented system, we computed time difference, direction, speed and curvature of each point along a stroke. The speed is calculated as $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them. The direction $d$ is calculated as the angle between two vectors and curvature is considered as the change in direction with respect to length i.e. $c= \Delta d/\Delta s$. Dominant points are characterized by lower speed values and higher curvature and direction values.

@@

As a modification to the system in \cite{mypaper}. After each iteration, of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, one of these points are removed based on the error caused by its removal.
	\item Each two successive segments are test to determine if they can be merged into a single segment. If merging the two segments have the same or smaller the error, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
\end{enumerate}

@@

To demonstrate the change in accuracy is due to the enhanced swarm optimization algorithm we tested the recognition system using two sets of features. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}) using the whole feature set. The second experiment was tested on both algorithms but only using half the feature set. Both experiments were compared with the Tradional PSO sketch recognition system presented in \cite{mypaper}.

Figures \ref{exp2} and \ref{exp1} show the recognition rate of each algorithm using the full feature set and half the features set respectively. The figures compares between the result that can be obtained using Traditional PSO by system in \cite{mypaper} and modifying the system using Enhanced PSO. The results show that the enhanced swarm optimization algorithm improves the segmentation of the system which leads to an increase in the recognition rate of the whole system by an average of $1.5\%$.  The increase in recognition rate is consistent in all swarm algorithms and in both experiments.   
    \begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment_92.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of the full feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO. The results show that Enhanced PSO increases the performance over the  Traditional PSO. }
   \label{exp2}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=3.3in]{Experiment_68.jpg}
  \caption[Recognition accuracy]%
  {The chart compares between recognition rate of half the feature set on both the traditional PSO system presented in \cite{mypaper} and Enhanced PSO. The results show that Enhanced PSO increases the recognition rate. }
   \label{exp1}
\end{figure}

@@

This paper used an Enhanced-PSO to modify a new sketch recognition using DPSO represented in \cite{mypaper}. Results show that enhancing the solutions that the \textit{PSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The final recognition improved by modifying the particle optimization algorithm without depending on the higher level features. The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on simple presentation set dataset. Results show that the enhanced PSO algorithm improves the segmentation on both algorithms which achieves better segmentation and final accuracy. The results proved that the Enhancement made on the PSO improves the efficiency of the Traditional PSO algorithm without being dependable on special objective function or application. The system achieved an average overall improvement of more than 1\% over Basic PSO.

@@

\bibliographystyle{elsarticle-num-names}
\bibliography{../../../neededfiles/Bibliographies/Mybibliography}
